---
title: "Economic History"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

```{r, include = F, echo = F}
require('readxl')
require('tidyverse')
require('formattable')
require('directlabels')
#install.packages('readxl')
#install.packages('tidyverse')
#install.packages('formattable')
#install.packages('directlabels')
library(readxl)
library(tidyverse)
library(formattable)
library(directlabels)
remove(list = ls())
#now I am going to read the excel sheet I need into the RMarkdown for graphing
IncomeTaxData <- read_excel('C:\\Users\\anden\\Dropbox\\Data Science\\web Tax Revenue Cyclicality\\IncomeTaxData_NA.xlsx', sheet="RExport")
IncomeTaxData2 <- read_excel('C:\\Users\\anden\\Dropbox\\Data Science\\web Tax Revenue Cyclicality\\IncomeTaxData_NA.xlsx', sheet="RExport1")

#IncomeTaxData$Year <- ymd(IncomeTaxData$Year,truncated=2L)
#IncomeTaxData$Year <- update(IncomeTaxData$Year, month = 6, mday = 30)

#??toc

```


# Tax Revenue Cyclicality

## Tax Revenue Cyclicality: The U.S. Example 1800-2020

By Nick Anderson

Dated 2022.01.07

### Abstract

This paper investigates the correlation between cyclicality and federal tax revenue in the United States. From the Funding Act of 1790, the United States has evolved alongside its tax system. After relying heavily on Customs Duties and Excise Taxes from 1790 to 1913, the United States turned to the Income Tax as its primary funding mechanism. Are these funding mechanisms procyclical or countercyclical? Can procyclical funding mechanisms reliably fund countercyclical fiscal policy? 

### Introduction

  Countercyclical fiscal policy was a cornerstone of the Keynesian revolution in economics. Keynes emphasized that, in economic downturns, governments who imposed austerity and cut spending only worsened their recessions by decreasing the level of aggregate demand. Thus, governments should respond to downturns with stimulus, spending against the cycle. To fund (or pay down the debt from) this stimulus, governments would run a surplus in upturns.
  The complication in the above policy prescription lies in the nature of the income tax. By the publication of Keynes’ General Theory in 1936, the income tax was a vital funding mechanism for the United States and United Kingdom, among many other countries. Because income taxes are earned by the government only when taxable incomes are earned, income taxes are inherently procyclical. This paper grew out of the author’s attempt to test the strength of the income tax’s cyclicality in U.S. history. 
  The starkest example of the income tax’s cyclicality occurred in the Great Depression. As unemployment increased to as high as 24% in 1933, U.S. income tax revenues fell over 60% (Figure 1). In the years of greatest revenue need, income tax revenues almost disappeared. 

```{r, echo = F}

#GDP <- 
IncomeTaxvs <- IncomeTaxData %>%
               select(Year,'Total Receipts') %>% 
               mutate(id='Total Receipts') %>%
               rename("type" = "id",
                      "value" = "Total Receipts",
                      "date" = "Year")

IncomeTaxvs1 <- IncomeTaxData %>%
               select(Year,'Not Income Tax') %>%
               mutate(id='Other Tax Revenue') %>%
               rename("type" = "id",
                      "value" = "Not Income Tax",
                      "date" = "Year")

IncomeTaxvs2 <- IncomeTaxData %>%
               select(Year,'Income Tax') %>% 
               mutate(id='Income Tax Revenue') %>%
               rename("type" = "id",
                      "value" = "Income Tax",
                      "date" = "Year")

IncomeTaxvs3 <- IncomeTaxData %>%
               select(Year,'Unemployment Rate') %>% 
               mutate(id='Unemployment Rate') %>%
               rename("type" = "id",
                      "value" = "Unemployment Rate",
                      "date" = "Year")

ITvs <- rbind(IncomeTaxvs,IncomeTaxvs1) 
ITvs <- rbind(ITvs,IncomeTaxvs2)
ITvs <- rbind(ITvs,IncomeTaxvs3)

ITvsCol <- ITvs %>%
           filter(type != 'Total Receipts') %>%
           filter(type != 'Unemployment Rate') %>%
           mutate(value = value/1000000000)

IncomeTaxvs3 <- IncomeTaxvs3 %>%
                mutate(mark=NA)


IncomeTaxvs3$mark[1] <- paste0(as.character(IncomeTaxvs3$value[1]*100), '%')
IncomeTaxvs3$mark[3] <- paste0(as.character(IncomeTaxvs3$value[3]*100), '%')
IncomeTaxvs3$mark[5] <- paste0(as.character(IncomeTaxvs3$value[5]*100), '%')
IncomeTaxvs3$mark[7] <- paste0(as.character(IncomeTaxvs3$value[7]*100), '%')
IncomeTaxvs3$mark[9] <- paste0(as.character(IncomeTaxvs3$value[9]*100), '%')
IncomeTaxvs3$mark[11] <- paste0(as.character(IncomeTaxvs3$value[11]*100), '%')
IncomeTaxvs3$mark[13] <- paste0(as.character(IncomeTaxvs3$value[13]*100), '%')
IncomeTaxvs3$mark[15] <- paste0(as.character(IncomeTaxvs3$value[15]*100), '%')

#ITvsCol2 <- rbind(ITvsCol,IncomeTaxvs3)
#str(IncomeTaxvs)

#ITvsCol$type <- relevel()

ggratio1 <- (max(IncomeTaxvs$value/1000000000))/max(IncomeTaxvs3$value)

ggIncomeTaxData <- ggplot(ITvsCol,aes(x=date,y=value,fill=type)) +
                   geom_col() +
                   scale_y_continuous(name="Dollars (in Billions)",
                                      breaks=seq(0,max(IncomeTaxvs$value/1000000000)+1.5,by=1),
                                      labels=scales::dollar_format(accuracy=.1),
                                      sec.axis =
                                      sec_axis(
                                        trans=~./max(IncomeTaxvs$value/1000000000)*max(IncomeTaxvs3$value),
                                                          name="% Unemployment Rate",
                                                          labels = scales::label_percent()),
                                      ) +
                   scale_x_continuous(name="Years of the Great Depression",
                                      breaks=seq(min(ITvsCol$date),max(ITvsCol$date),by=2),
                                      ) +
                   ggtitle("Figure 1: U.S. Federal Government Revenue from 1927-1941, 
                           with Unemployment Rate") +
                   guides(fill=guide_legend(title='')) +
                   scale_fill_manual(values=c("#640000", "#00B0F0",'black')) +
                   theme(legend.position="right",
                         panel.grid.major = element_blank(), 
                         panel.grid.minor = element_blank(),
                         panel.background = element_blank(), 
                         axis.line = element_line(colour = "black"), 
                         ) +
                   geom_line(data=IncomeTaxvs3,
                             aes(y=value*ggratio1),
                             size=2,
                             ) +
                   geom_label(data=IncomeTaxvs3,
                              aes(y = value*ggratio1+.4,
                                  label = mark,), 
                              nudge_x = 0.35, size = 3,
                              fill = 'gray',
                             ) 
                   

ggIncomeTaxData

#?geom_dl
#?scales::percent
#?scales::label_percent()
#?geom_hline
#?geom_label

#panel.border = element_rect(color = "black", fill=NA, size = 3),
# panel.grid.major.y = element_line(color = "black", size = 0.1, linetype = 2)


```

Source: Lebergott 1957 (he was with the NBER) and the BLS (My Data)LINK

The Great Depression remains the severest example of economic depression available for examination. The author emphasizes this example not as conclusive proof, but as a demonstration of income tax revenue procyclicality.

The ‘business cycle’ is not easy to define, so this paper examines two aspects of cyclicality: Gross Domestic Product (as a measure of incomes) and the Unemployment Rate (as a measure of employment). The graph above replicated with Nominal GDP instead of the unemployment rate tell the similar stories from different perspectives (Figure 2). When aggregate demand contracts, the income tax contracts as well.


```{r, echo = F}

ITvsCol2 <- rbind(ITvsCol,IncomeTaxData2)

ITvsCol2a <- ITvsCol2 %>%
            filter(type != 'NGDP' & type != 'Population')

ITvsCol2b <- ITvsCol2 %>%
            filter(type == 'NGDP')

#ITvsCol2b[1] <- ITvsCol2b[1]/10

ggIncomeTaxData2 <- ggplot(ITvsCol2a,aes(x=date,y=value,fill=type)) +
                   geom_col() +
                   scale_y_continuous(name="Tax Revenue in $B",
                                      breaks=seq(0,max(IncomeTaxvs$value/1000000000)+1.5,by=1),
                                      labels=scales::dollar_format(accuracy=.1),
                                      sec.axis =
                                      sec_axis(
                                        trans=~./max(IncomeTaxvs$value)*max(ITvsCol2b$value),
                                                          name="Nominal GDP in $B",
                                                          labels=scales::dollar_format(accuracy=.1)),
                                      ) +
                   scale_x_continuous(name="Years of the Great Depression",
                                      breaks=seq(min(ITvsCol$date),max(ITvsCol$date),by=2),
                                      ) +
                   ggtitle("Figure 1: U.S. Federal Government Revenue from 1927-1941, 
                           with Unemployment Rate") +
                   guides(fill=guide_legend(title='')) +
                   scale_fill_manual(values=c("#640000", "#00B0F0",'black')) +
                   theme(legend.position="right",
                         panel.grid.major = element_blank(), 
                         panel.grid.minor = element_blank(),
                         panel.background = element_blank(), 
                         axis.line = element_line(colour = "black"), 
                         ) +
                   geom_line(data=ITvsCol2b,
                             aes(y=value/max(IncomeTaxvs$value)),
                             size=2, 
                             color = '#00B0F0')

ggIncomeTaxData2

```


```{r, echo = F}
# d <- data.frame(
#   y=c(0.1, 0.2, 0.7),
#   cat = factor(c('No', 'Yes', 'NA'), levels = c('NA', 'Yes', 'No')))
# 
# # Original order
# p1 <- ggplot(d, aes(x=1, y=y, fill=cat)) +
#   geom_bar(stat='identity')
# 
# # Change order of rows
# p2 <- ggplot(d[c(2, 3, 1), ], aes(x=1, y=y, fill=cat)) +
#   geom_bar(stat='identity')
# 
# # Change order of levels
# d$cat2 <- relevel(d$cat, 'Yes')
# p3 <- ggplot(d, aes(x=1, y=y, fill=cat2)) +
#   geom_bar(stat='identity') 
# 
# grid.arrange(p1, p2, p3, ncol=3)
```


# Second Project

```{r, echo = F}
samples <- rnorm(100, mean=0, sd=1)
hist(samples)
```


# IRS Data


2018 | 7 | 23 Last compiled: `r Sys.Date()`





<style type="text/css">
.title {
  display: none;
}

.navbar-default {
    background-color: #640000 !important;
    
}

<!-- .navbar-nav { -->
<!--   color:white !important; -->

<!-- } -->

.navbar .navbar-brand, .navbar .navbar-nav .nav-links  {
  color: white;
}

.navbar-default .navbar-nav>li>a {
    color: white !important;
    font-weight: 150;
}

.navbar-nav > .active > a { 
    background-color: #006400 !important;
    font-weight: 175;
}

#getting-started img {
  margin-right: 10px;
  margin-left: 5px;
}


</style>

<div class="row" style="padding-top: 30px;">
<div class="col-sm-6">
