---
title: "Economic History"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Tax Revenue Cyclicality: The U.S. Example
By Nick Anderson
Dated 2022.01.07

### Abstract

This paper investigates the correlation between cyclicality and federal tax revenue in the United States. From the Funding Act of 1790, the United States has evolved alongside its tax system. After relying heavily on Customs Duties and Excise Taxes from 1790 to 1913, the United States turned to the Income Tax as its primary funding mechanism. Are these funding mechanisms procyclical or countercyclical? Can procyclical funding mechanisms reliably fund countercyclical fiscal policy? 

### Introduction

  Countercyclical fiscal policy was a cornerstone of the Keynesian revolution in economics. Keynes emphasized that, in economic downturns, governments who imposed austerity and cut spending only worsened their recessions by decreasing the level of aggregate demand. Thus, governments should respond to downturns with stimulus, spending against the cycle. To fund (or pay down the debt from) this stimulus, governments would run a surplus in upturns.
  The complication in the above policy prescription lies in the nature of the income tax. By the publication of Keynes’ General Theory in 1936, the income tax was a vital funding mechanism for the United States and United Kingdom, among many other countries. Because income taxes are earned by the government only when taxable incomes are earned, income taxes are inherently procyclical. This paper grew out of the author’s attempt to test the strength of the income tax’s cyclicality in U.S. history. 
  The starkest example of the income tax’s cyclicality occurred in the Great Depression. As unemployment increased to as high as 24% in 1933, U.S. income tax revenues fell over 60% (Figure 1). In the years of greatest revenue need, income tax revenues almost disappeared. 



2018 | 7 | 23 Last compiled: `r Sys.Date()`

##Abstract
asd

###Abstract
asd

## Abstract
asd

### Abstract
asd


## Second level header
### Third Level Header
#Project 2



I'm writing this tutorial going from the top down. And, this is how it will be printed. So, notice the second post is second in the list. If you want your most recent post to be at the top, then make a new post starting at the top. If you want the oldest first, do, then keep adding to the bottom

# Adding R stuff

So far this is just a blog where you can write in plain text and serve your writing to a webpage. One of the main purposes of this lab journal is to record your progress learning R. The reason I am asking you to use this process is because you can both make a website, and a lab journal, and learn R all in R-studio. This makes everything really convenient and in the sam place. 

So, let's say you are learning how to make a histogram in R. For example, maybe you want to sample 100 numbers from a normal distribution with mean = 0, and standard deviation =1, and then you want to plot a histogram. You can do this right here by using an r code block, like this:

```{r}
samples <- rnorm(100, mean=0, sd=1)
hist(samples)
```

2018 | 7 | 23 Last compiled: `r Sys.Date()`
